---
title: "owner-analysis"
author: "Julek Chawarski"
format: html
---

# load libraries

```{r, warnings = FALSE, echo = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

# read in the data

```{r}
lobster_abundance <- read_csv(here("data/Lobster_Abundance_All_Years_20220829.csv"))
```

# convert missing values to NA

```{r}
lobster_abundance <- lobster_abundance %>% 
  mutate(SIZE_MM = na_if(SIZE_MM, -99999))


```

# remove observations from Naples Reef
```{r}
lobster_abundance_noNAPL <- lobster_abundance %>%
  filter(!SITE %in% "NAPL")
```

# create and store a subset from Arroyo Quemado and with carapace length > 70 mm

```{r}
lobster_abundance_AQUE <- lobster_abundance %>% 
  filter(SITE %in% "AQUE") %>%
  filter(SIZE_MM > 70)


hist(lobster_abundance_AQUE$SIZE_MM)
```

# find max carapace length, group by site and month

```{r}
lobster_size <- lobster_abundance %>% 
  group_by(SITE, MONTH) %>%
  summarise(max_size = max(SIZE_MM, na.rm = T))
```

## Exercise: Data Visualization

# bar plots of maximum lobster sizes by size and month

```{r}

lobster_size %>% ggplot(aes(x=SITE, y= max_size, fill=SITE)) + geom_col() +
  facet_grid(~MONTH) + theme_bw() + theme(legend.position = "none")
```

# lobster carapace size distributions by SITE

```{r}

plot3 <- 
lobster_abundance %>% ggplot(aes(x=SIZE_MM, fill=SITE)) + 
  geom_histogram(color="black") + xlab("Size (mm)") + 
  labs(title = "Lobster Carapace Size Distributions by Site, 2012-2022") + 
  facet_wrap(~SITE) + theme_minimal() +  theme(legend.position = "none")

ggsave(here("figs/Lobster_carapace_sizedist_site.png"), plot = plot3, height=4, width = 6, dpi=300)


```

# total count of lobsters by SITE and YEAR

```{r}
lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>%
  summarise(sum = sum(COUNT))

plot2 <- 
lobsters_summarize %>% ggplot(aes(x=YEAR, y=sum, color=SITE, linetype = SITE)) + 
  geom_point() + 
  scale_x_continuous(breaks = c(2012:2022)) + 
  geom_line() + 
  labs(x = "Year", y = "# lobsters", title= "Total Count of Lobsters across sites and years, 2021-2022") + 
  theme_minimal() 

ggsave(here("figs/Total_lobster_count_sitexyear.png"), plot = plot2, height=3, width = 6, dpi=300)

```


# Lobster size analysis

```{r}

lobster_size_lrg <- lobster_abundance %>%
  filter(YEAR %in% c(2019,2020,2021)) %>%
  mutate(SIZE_BIN = if_else(SIZE_MM <= 70, "small", "large")) %>%
  group_by(SITE, SIZE_BIN) %>%
  summarise(num = sum(COUNT)) %>%
  drop_na()

cols = c("cadetblue", "chocolate3")

plot1 <- 
ggplot(data = lobster_size_lrg, aes(x = SITE, y = num, fill = SIZE_BIN)) +
    geom_col(position = "dodge", color="black") + 
  labs(title = "Lobster size groups by Site, all years", y= "# Lobsters") + 
  scale_fill_manual(values = cols, labels = c("Large (>= 70 mm)", "Small (< 70 mm)"), name = "Size Groups") + 
  theme_minimal() +
  labs(legend.title = "Size group")

ggsave(here("figs/Lobster_size_groups_plot.png"), plot = plot1, height=3, width = 5, dpi=300)

```



